Search for WIMP Inelastic Scattering Off Xenon Nuclei With Xenon100 Data

### 2017-01-16 Inelastic/intro.tex
Laura's modifications to intro.tex

\section{\label{sec:level1} Introduction}
Astrophysical evidence indicates that the dominant mass fraction of our Universe consists of some yet unknown form
of dark matter. Well motivated models predict Dark Matter in the form of Weakly Interacting Massive Particles (WIMPs),
hypothesis which is currently being tested by several direct and indirect detection experiment.

Most of direct detection searches focuses on  elastic scattering of dark matter particles off nuclei.
In this analysis instead we explore an inelastic scattering process, we consider the $^{129}\text{Xe}$ isotope being excited to a low-lying state
with subsequent prompt de-excitation via the emission a photon. This isotope is an excellent target
since its abundance in natural xenon is of 26.4\% and a relatively low energy is necessary to excite its $3/2+$ state above the $1/2+$
spin ground state.
Inelastic WIMP-nucleus scattering in xenon is complementary to elastic scattering for spin-dependent interactions,
the former dominates the integrated rate above  $\simeq10$~keV of energy deposition.
Furthermore, in the case of dark matter detection, this channel can be employed to asses whether the nature of the fundamental interaction
is spin-dependent or not.

\section{Xenon100 Detector}
The Xenon100 experiment is a  dual phase liquid xenon TPC. For a given interaction in the liquid target this type of detector produces two separated signals,
one proportional to the prompt scintillation (S1) the other to ionization (S2).

To add: sentences about detector stability, science run data used, Ly and Y measurements used.

>>>>

\section{\label{sec:level1} Introduction}
Astrophysical and cosmological evidence indicates that the dominant mass fraction of our Universe consists of some yet unknown
form of dark, or invisible matter. The dark matter could be made of new, yet undiscovered particles, and well-motivated
theoretical models going beyond the Standard Model of particle physics predict the existence of Weakly Interacting Massive
Particles (WIMPs), which are natural candidates for dark matter. This hypothesis is currently being tested by several direct
and indirect detection experiments, as well as at the LHC~\cite{Bertone:2004pz,Baudis:2016qwx}.

Most direct detection searches focus on elastic scattering of galactic dark matter particles off nuclei, where the keV-scale 
nuclear recoil energy is to be detected~\cite{Baudis:2012ig,Baudis:2015mpa,Undagoitia:2015gya}. In this work, the 
alternative process of inelastic scattering is explored, where a WIMP-nucleus scattering induces a transition to a low-lying 
excited nuclear state. The experimental signature is a nuclear recoil detected together with the prompt de-excitation 
photon~\cite{Ellis:1988nb}.

We consider the $^{129}\text{Xe}$ isotope, which has an abundance of 26.4\% in natural xenon, and a lowest-lying 
3/2$^{+}$ state at 36.6\,keV above the 1/2$^+$ gound state. The electromnagnetic nuclear decay has a half-life of 0.97\,s. 
The sigantures of inelastic scattering in xenon have been studied in detail in~\cite{Baudis:2013bba}. It was found that this 
channel is complementary to spin-dependent, elastic scattering, dominating the integrated rates above $\simeq10$\,keV energy 
depositions. In addition, in case of a positive signal, the observation of inelastic scattering would provide a clear 
indication of the spin-dependent nature of the fundamental interaction.

\section{The XENON100 Detector}

The XENON100 experiment operated a dual-phase (liquid and gas) xenon time projection chamber (TPC) at the Laboratori Nazionali 
del Gran Sasso (LNGS) in Italy. It contained 161\,kg of xenon in total,  with  62\,kg in the active region of the TPC. These 
were monitored by 242 low-radioactivity, UV-sensitive photomultiplier tubes (PMTs) arranged in two arrays, one in the liquid 
and one in the gas. The PMTs detected the prompt scintillation (S1) and the delayed, proportional scintillation signal (S2) 
created by a particle interacting in the active TPC region. The S2-signal is generated due to ionisation electrons, drifted 
and extracted into the gas phase, where the proportional scintillation, or electroluminiscence, is produced. These photons 
carry the $x-y$ information of the interaction site, while the $z-$information came from the drift time measurement. The TPC 
thus yielded a three-dimensional event localisation, enabling to reject the majority of background events via fiducial volume 
cuts, as well a multiple interactions, which are unlikely due to dark matter particles~\cite{Aprile:2011dd}. The ratio of 
S2/S1 provided the basis for distinguishing between nuclear recoils (NRs), as induced by fast neutrons and expected from 
elastic WIMP-nucleus scatters, and electronic recoils (ERs) produced by $\beta$ and $\gamma$-rays. As we shall see, the 
inelastic scattering signature will be a superposition of an ER and a NR, due to the low lifetime of the excited state and 
the short mean free path ($\sim$0.15\,mm) of the 39.6\,keV photon.

XENON100 has acquired science data between 2008-2015, and has set competitive constraints on spin-
independent~\cite{Aprile:2012nq,Aprile:2016swn} and spin-dependent~\cite{Aprile:2013doa,Aprile:2016swn} elastic WIMP-nucleus 
scatters, as well as on axions/ALPs~\cite{Aprile:2014eoa} and leptophilic dark matter 
models~\cite{Aprile:2015ade,Aprile:2015ibr,Aprile:2017yea}.
